---
title: "Smolt data import Report"
author: "Anders Kagervall"
date: '`r Sys.Date()`'
output:
  html_document: default
params:
  smoltreg: NA
  origName: "name missing"
---


```{r initialize, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("config.R", encoding = "UTF-8")
source("functions.R", encoding = "UTF-8")

library(dplyr, quietly = TRUE)
smoltreg <- params$smoltreg
metadata <- read_meta(smoltreg)
fishdata <- read_fish(smoltreg, dummy_tags = metadata$dummy_tags)
allowed_species <- read.table("allowed_species.txt", sep="\t", encoding="UTF-8")[,1]
river <- metadata$river
year <- format(metadata$startdate, "%Y")
```

## Smolt trap `r river` `r year`

This report contains an automated report of things that *may* indicate errors or problems
in the data file *`r params$origName`*. When the script detects a problem it is up to
you to fix it or decide that it actually is OK before doing further analyses.


### Species

After removal of empty rows from the data file *`r params$origName`* contains `r nrow(fishdata)` rows with capture of a fish. **Check that the numbers seems reasonable.**

The number of fish per species:
```{r speciestable}
sort(table(fishdata[fishdata$event != RECAPTURED,]$species), decreasing =  TRUE)
```

Check for unknown species in input.
```{r speciestable2}
as.data.frame(fishdata[!fishdata$species %in% allowed_species &
                         fishdata$event != RECAPTURED,])
```


### Data format

Checking data format in input files. If any test say **FIX IT** you must FIX IT before 
running xlsx2sote.R or you will not get correct data.



```{r datacheck_format, echo=FALSE, results='asis'}
##
LENGTH_WEIGHT_OK <- TRUE # First assume check will be OK
cat(paste0("**Checking fish data format in selected columns read from *", params$origName, "*.**\n\n"))
##
#need.date <- fishdata[!is.na(fishdata$pittag),]$date_time # all rows with pittag must hav valid date
if (any(is.na(as.POSIXct(fishdata$date_time)))) {
  cat("+ Column *date_time* contains data that does not convert to a date, **FIX IT**.\n")
} else {
  cat("+ Column *date_time* looks OK.\n")
}
num.columns <- c("length", "weight")
for (name in num.columns ) {
  if (is.possible.numeric(fishdata[ ,name])) {
    cat(paste("+ Column *", name, "* is numeric, good.\n", sep = "")) 
  } else {
    cat(paste("+ Column *", name, "* is **NOT** numeric, **FIX IT**.\n", sep = ""))
    LENGTH_WEIGHT_OK <- FALSE
  }
}
##
# cat("\n")
# cat(paste0("**Checking environmental data format in selected columns read from *", envfile, "*.**\n\n"))
# ##
# if (any(is.na(as.POSIXct(envdata$date)))) {
#   cat("+ Column *date_time* contains data that does not convert to a date, **FIX IT**.\n")
# } else {
#   cat("+ Column *date_time* looks OK.\n")
# }
# num.columns <- c("w_level", "w_temp")
# for (name in num.columns ) {
#   if (is.possible.numeric(envdata[ ,name])) {
#     cat(paste("+ Column *", name, "* is numeric, good.\n", sep = "")) 
#   } else {
#     cat(paste("+ Column *", name, "* is **NOT** numeric, **FIX IT**.\n", sep = ""))
#   }
# }
```

### Smolt status
The column smoltstat can only contain values S0, S1, S2 or S3 (or be blank). **If a table with fish is printed below, fix the smoltstat and re-run.**
```{r smoltstat, echo = FALSE, results='asis'}

stab <- fishdata %>%
  filter(!is.na(smoltstat)) %>%
  filter(!(smoltstat %in% c('S0', 'S1', 'S2', 'S3')))

if (nrow(stab) > 0) {
  knitr::kable(stab,
               caption = "Fish with unknown smoltstat.")
} else {
  cat("No fish with unknown smoltstat. :-)")
}
```

### Genid
Check that no genid are duplicated. **If a table with fish is printed below, fix the smoltstat and re-run.**

```{r genid, echo = FALSE, results='asis'}
dups <- duplicated(fishdata$genid) & !is.na(fishdata$genid)
if (any(dups)) {
  knitr::kable(fishdata[dups,],
               caption = "Fish with duplicated genid.")
  
} else {
   cat("No fish with duplicated genid. :-)") 
}
```

### Event types

The imported rows are classified as an event type that can be either: a caught (but not marked) fish, a fish that gets marked or a recaptured marked fish. The event type should not be unknown for any rows. **Check that the number of unknown events is zero (in some cases unknowns can be OK) and that the other numbers seems reasonable.**

```{r eventtable}
events <- factor(fishdata$event, c(0:4), labels = c("unknown", "caught",
                                                    "marked", "recaptured", "removed"))
table(events)
```



All fish with event `r MARKED`(=MARKED) or  `r RECAPTURED`(=RECAPTURED) should have a pittag registered. **If a tables with fish is printed, check if the event type is wrong.**

```{r eventhavepittag, echo = FALSE, results='asis'}

etab <- fishdata %>%
  filter(event %in% c(MARKED, RECAPTURED)) %>%
  filter(is.na(pittag))
if (nrow(etab) > 0) {
  knitr::kable(etab,
               caption = "Marked or recaptured fish without pittag. Probably wrong event type")
} else {
  cat("No marked or recaptured fish without pittag. :-)")
}
```



### Duplicates (fix errors under Event type first)

A pittag should only be used one time in a marking event and one or zero times in a recapture event. We check this by searching for duplicate pittag id in the dataset.
**If tables with pit tags are printed, check the reason for the duplicated pittags.**

```{r dups, echo = FALSE, results='asis'}
marked <- fishdata[fishdata$event == MARKED, ] # all rows with marking event
recap <- fishdata[fishdata$event == RECAPTURED, ] # all rows with recapture event 
marked.dups.IDs <- marked[duplicated(marked$pittag), ]$pittag # all 
recap.dups.IDs <- recap[duplicated(recap$pittag), ]$pittag
mtab <- marked %>% filter(pittag %in% marked.dups.IDs) %>% arrange(pittag)
rtab <- recap %>% filter(pittag %in% recap.dups.IDs) %>% arrange(pittag)
if (nrow(mtab) > 0) {
  knitr::kable(mtab,
               caption = "Duplicated pittags in marked fish") # Show duplicated pittags in marked fish
} else {
  cat("No duplicates found in marked fish. :-)")
}
```

```{r dups2, echo=FALSE, results = 'asis'}
#
if (nrow(rtab) > 0) {
  knitr::kable(rtab,
               caption = "Duplicated pittags in recaptured fish")  # Show duplicated pittags in recaptured fish
} else {
  cat("No duplicates found  in recaptured fish. :-)")
}  
```

### Not marked recaptures

A fish recaptured should always have a mark event before the recapture. 
**If tables with pit tags are printed, check the reason for the recapture without marking event.**

```{r notmarked, echo = FALSE, results = 'asis'}
r.ids <- unique(recap$pittag)
m.ids <- unique(marked$pittag)
recap.not.marked <- r.ids[!(r.ids %in% m.ids)]
not.marked.tab <- fishdata[fishdata$pittag %in% recap.not.marked, ]
if (nrow(not.marked.tab) > 0) {
  knitr::kable(not.marked.tab, caption = "Recaptured but no marking event")
} else {
  cat("No unmarked recaptures found. :-)")
}
```


### Species error
The species recorded for a marked fisk must be consistent within the data file. **Check the reason if different species are registered for any pittag.**

```{r specieserror, echo = FALSE, results = 'asis'}
sp.err <- fishdata %>%
  filter(!is.na(pittag)) %>%
  filter(!is.na(species)) %>%
  group_by(pittag) %>%
  mutate(scount = length(unique(species))) %>%
  filter(scount > 1)
sp.err.ids <- unique(sp.err$pittag)
sp.err.tab <- fishdata[fishdata$pittag %in% sp.err.ids,] %>% arrange(pittag)
if (nrow(sp.err.tab) > 0) {
  knitr::kable(sp.err.tab, caption = "Pittag with more than one species registered")
} else {
  cat("No species errors found. :-)")
}
```

### Fulton

Calculate mean fulton(if possible):
```{r meanfulton}
if (LENGTH_WEIGHT_OK) { # Only execute if length and weight are numeric
  f <- fishdata %>%
    filter(species %in% c('Lax', 'Öring'), length < 400) %>%
    mutate(k = fulton(weight, length))
  with(f, tapply(k, species, mean, na.rm=TRUE))
}
```

Check that all measured fish have a fulton between `r min_k` and `r max_k`.

```{r datacheck_fulton, echo=FALSE, results='asis'}
#min_k <- params$min_k
#max_k <- params$max_k
if (LENGTH_WEIGHT_OK) { # Only execute if length and weight are numeric
  bad.condition <- fishdata %>% 
    filter(!is.na(pittag) & !is.na(length) & !is.na(weight)) %>%
    filter(length < 400) %>%
    mutate(k = fulton(weight, length)) %>%
    filter(k < min_k | k > max_k)
  if (nrow(bad.condition) > 0) {
    knitr::kable(bad.condition,
                 caption = "Fulton outside limits. Check length and weight")
  } else {
    cat("All fish have condition factors within limits. :-)") }
  } else {
    cat('Can not compute fulton. *FIX* length and weight data format.')
  }

```

### Metadata

Below is av printout of data found in sheets *Metadata* and *Metadata2*. *Metadata* should be filled by the personel running the trap and *Metadata2* by SLU.
```{r metadata}
metadata
```

